---
title: "R Notebook"
output: html_notebook
---


```{r, warnings=FALSE, message=FALSE}
library(tidyverse)
library(marginaleffects)
library(fixest)
library(gam)
library(splines)
library(ggplot2)
library(ggpubr)
library(gridExtra)
library(useful)
library(purrr)
library(patchwork)
library(rms)
library(tidycensus)
library(tigris)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(did)
library(haven)
library(bea.R)
library(lpdid)
library(DIDmultiplegtDYN)
library(ggpattern)
library(countrycode)
library(stringr)
library(ggpattern)
library(scales)
library(RColorBrewer)
library(ggpubr)
library(splp)
library(MatchIt)
library(cobalt)
library(cowplot)

options(tigris_use_cache = TRUE)
setFixest_notes(FALSE)
source("../emileRegs.R")
```





# Data  

```{r}
myblue  = "#1f78b4"   # blue
myred  = "#e31a1c"
myyellow = "orange"  # orange
myblack = "black"  # light blue
mygreen = "forestgreen"  # green
```



```{r}
# Also load full panel for comparison
pwt  <- read_stata("../raw_data/pwt_clean.dta")
tcs  <- read_stata("../raw_data/ibtracs_clean.dta")
data_full <- pwt %>% left_join(tcs, by = c("year", "countrycode"))

knot_to_ms <- 0.514444
data <- data_full %>%
  mutate(
    maxwind  = replace_na(max_ann_wind_i_nob, 0) * knot_to_ms,
    loggdp   = ln_real_gdp_usd_pc,
    gdp_diff = growth_real_gdp_usd_pc,
    year2    = year^2,
    period   = ifelse(year <= 1990, "Pre-1990", "Post-1990")
  ) %>%
  filter(year >= 1969) %>%
  mutate(
    maxwind_p95 = quantile(maxwind[year >= 1970 & year <= 2014], 0.95, na.rm = TRUE),
    maxwind_95 = as.integer(maxwind_p95 > 0 & maxwind >= maxwind_p95),
    cat1_plus = as.integer(maxwind >= 32.7),  # Cat 1+ threshold in m/s
    cat2_plus = as.integer(maxwind >= 49.4),  # Cat 2+ threshold in m/s
    cat3_plus = as.integer(maxwind >= 70.0),   # Cat 3+ threshold in m/s
    region = countrycode(countrycode, "iso3c", "region"),
    region = ifelse(is.na(region), "Other", region),
    continent = countrycode(countrycode, "iso3c", "continent"),
    continent = ifelse(is.na(continent), "Other", continent)
  )

country_gdp <- data %>%
  filter(!is.na(loggdp)) %>%
  group_by(countrycode) %>%
  summarise(
    mean_loggdp = mean(loggdp, na.rm = TRUE),
    mean_gdp_pc = exp(mean(loggdp, na.rm = TRUE)),
    n_shocks_95 = sum(maxwind_95, na.rm = TRUE),
    total_wind  = sum(maxwind, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    income_quartile = ntile(mean_loggdp, 4),
    income_quartile_label = factor(
      income_quartile,
      levels = 1:4,
      labels = c("Q1 (Poorest)", "Q2", "Q3", "Q4 (Richest)")
    )
  )

data <- data %>%
  left_join(
    country_gdp %>% select(countrycode, income_quartile, income_quartile_label,
                            mean_loggdp, mean_gdp_pc),
    by = "countrycode"
  )

```{r}

data <- data %>%
  group_by(countrycode) %>%
  mutate(
    cat1_plus_lag1 = lag(cat1_plus, 1, default = 0),
    cat1_plus_lag2 = lag(cat1_plus, 2, default = 0),
    cat1_plus_lag3 = lag(cat1_plus, 3, default = 0),
    cat1_plus_lag4 = lag(cat1_plus, 4, default = 0),
    cat1_plus_lag5 = lag(cat1_plus, 5, default = 0)
  ) %>%
  ungroup()


ever_treated_countries <- data %>%
  filter(maxwind > 0) %>%
  pull(countrycode) %>%
  unique() %>%
  sort()

d_treated <- data %>%
  filter(countrycode %in% ever_treated_countries)

```




# Exposure

```{r}
outcome  <- "loggdp"
horizon  <- 10
make_controls <- function(shock_var) {
  paste0("l(gdp_diff,1:2) + l(", shock_var, ",1:2)")
}
fe       <- "countrycode[year] + countrycode[year2] + region^year"
panel_id <- c("countrycode", "year")
vcov_fm  <- DK ~ year
```


```{r}


world <- ne_countries(scale = "medium", returnclass = "sf")

# Country-level tercile lookup (one row per country)
tercile_map <- data %>%
  distinct(countrycode, cat1_tercile)

map_df <- world %>%
  left_join(tercile_map, by = c("adm0_a3" = "countrycode")) %>%

  mutate(
    cat1_tercile = as.character(cat1_tercile),
    cat1_tercile = case_when(
      is.na(cat1_tercile) ~ "No data",
      TRUE                ~ cat1_tercile
    ),
    cat1_tercile = factor(cat1_tercile,
                          levels = c("Never treated",
                                     "Low exposure",
                                     "Mid exposure",
                                     "High exposure",
                                     "No data"))
  ) %>% 
    #remove na
  filter(cat1_tercile != "No data")

map_colors <- c(
  "Never treated"      = "#d9d9d9",
  "Low exposure"  = "#9ecae1",
  "Mid exposure"  = "#f6a623",
  "High exposure" = "#c0392b",
  "No data"            = "midnightblue"
)

fig_map <- ggplot(map_df) +
  geom_sf(aes(fill = cat1_tercile), color = "white", linewidth = 0.15) +
  coord_sf(crs = "+proj=robin") +
  scale_fill_manual(
    values = map_colors,
    guide = guide_legend(
      title          = "Cat 1+ exposure",
      title.position = "top",
      nrow           = 2,
      keywidth       = unit(1.2, "cm"),
      keyheight      = unit(0.4, "cm")
    )
  ) +
  theme_void(base_size = 13) +
  theme(
    legend.position   = "bottom",
    legend.title      = element_text(face = "bold", size = 11),
    legend.text       = element_text(size = 10),
    plot.background   = element_rect(fill = "white", color = NA),
    panel.background  = element_rect(fill = "white", color = NA)
  )

ggsave("figures/map_cat1_terciles.png",
       fig_map, width = 5, height = 5, dpi = 300)

fig_map
```


```{r}
# ------------------------------------------------------------------
# 8. IRFs: Cat1+ treatment × historical exposure terciles
# ------------------------------------------------------------------

# Ever-treated: total Cat1+ events per country over the full sample
data <- data %>%
  group_by(countrycode) %>%
  mutate(country_cat1_count = sum(cat1_plus, na.rm = TRUE)) %>%
  ungroup()

# Terciles only among ever-treated countries (count > 0)
tercile_breaks <- data %>%
  filter(country_cat1_count > 0) %>%
  distinct(countrycode, country_cat1_count) %>%
  pull(country_cat1_count) %>%
  quantile(c(1/3, 2/3), na.rm = TRUE)

data <- data %>%
  mutate(
    cat1_tercile = case_when(
      country_cat1_count == 0                          ~ "Never treated",
      country_cat1_count <= tercile_breaks[1]          ~ "Low exposure",
      country_cat1_count <= tercile_breaks[2]          ~ "Mid exposure",
      country_cat1_count >  tercile_breaks[2]          ~ "High exposure"
    ),
    cat1_tercile = factor(cat1_tercile,
                          levels = c("Never treated",
                                     "Low exposure",
                                     "Mid exposure",
                                     "High exposure"))
  )

cat("\n--- Country Cat1+ exposure distribution ---\n")
data %>%
  distinct(countrycode, cat1_tercile, country_cat1_count) %>%
  count(cat1_tercile) %>%
  print()

tercile_colors <- c(
  "Never treated"      = "grey60",
  "Low exposure"  = "steelblue",
  "Mid exposure"  = "goldenrod",
  "High exposure" = "firebrick"
)

irfs_cat1_tercile <- lp_panel_inter(
  data         = data,
  outcome      = outcome,
  main_var     = "cat1_plus",
  interact_var = "cat1_tercile",
  controls     = make_controls("cat1_plus"),
  horizon      = horizon,
  fe           = fe,
  panel_id     = panel_id,
  vcov_formula = vcov_fm
) %>%
  mutate(
    exposure = factor(category,
                      levels = c("Never treated",
                                 "Low exposure",
                                 "Mid exposure",
                                 "High exposure"))
  )

# Print at h = 5
cat("\n--- Cat1+ × exposure tercile, h=5 ---\n")
irfs_cat1_tercile %>%
  filter(horizon == 5) %>%
  select(exposure, irf_mean, se, irf_down, irf_up) %>%
  arrange(exposure) %>%
  as.data.frame() %>%
  print(digits = 4, row.names = FALSE)

fig_cat1_tercile <- irfs_cat1_tercile %>%
  ggplot(aes(x = horizon, y = irf_mean,
             color = exposure, fill = exposure)) +
  geom_hline(yintercept = 0, linetype = "dashed",
             linewidth = 0.55, color = "grey40") +
  geom_ribbon(aes(ymin = irf_down, ymax = irf_up),
              alpha = 0.13, color = NA) +
  geom_line(linewidth = 1.05) +
  scale_color_manual(values = tercile_colors) +
  scale_fill_manual(values  = tercile_colors) +
  scale_x_continuous(breaks = seq(0, horizon, 2)) +
  scale_y_continuous(labels = pct_labels) +
  labs(x     = "Horizon (years)",
       y     = "GDP response",
       color = "Country Cat1+ exposure",
       fill  = "Country Cat1+ exposure") +
  irf_theme +
  theme(legend.position = "top") +
  #remove legend
  theme(legend.position = "none")


ggsave("figures/map_cat1_terciles.png",
       fig_map, width = 7, height = 5, dpi = 300)

ggsave("figures/irf_cat1_exposure_tercile.png",
       fig_cat1_tercile, width = 7, height = 5, dpi = 300)

fig_combined <- fig_map + fig_cat1_tercile + plot_layout(widths = c(1, 1))


ggsave("figures/combined_cat1_tercile.png",
       fig_combined, width = 10, height = 5, dpi = 300)
```


```{r}
# ------------------------------------------------------------------
# IRFs: Cat1+ treatment × country mean maxwind (continuous) terciles
# ------------------------------------------------------------------

# Country mean of continuous maxwind (all years, regardless of threshold)
data <- data %>%
  group_by(countrycode) %>%
  mutate(country_maxwind_mean = mean(maxwind, na.rm = TRUE)) %>%
  ungroup()

# Terciles only among countries with any positive maxwind
tercile_breaks_mw <- data %>%
  filter(country_maxwind_mean > 1) %>%
  distinct(countrycode, country_maxwind_mean) %>%
  pull(country_maxwind_mean) %>%
  quantile(c(1/3, 2/3), na.rm = TRUE)

data <- data %>%
  mutate(
    maxwind_tercile = case_when(
      country_maxwind_mean == 0                          ~ "Never exposed",
      country_maxwind_mean <= tercile_breaks_mw[1]      ~ "Low exposure",
      country_maxwind_mean <= tercile_breaks_mw[2]      ~ "Mid exposure",
      country_maxwind_mean >  tercile_breaks_mw[2]      ~ "High exposure"
    ),
    maxwind_tercile = factor(maxwind_tercile,
                             levels = c("Never exposed",
                                        "Low exposure",
                                        "Mid exposure",
                                        "High exposure"))
  )

cat("\n--- Country maxwind (continuous) tercile distribution ---\n")
data %>%
  distinct(countrycode, maxwind_tercile, country_maxwind_mean) %>%
  count(maxwind_tercile) %>%
  print()

# Sanity check: mean maxwind by tercile
data %>%
  distinct(countrycode, maxwind_tercile, country_maxwind_mean) %>%
  group_by(maxwind_tercile) %>%
  summarise(
    n         = n(),
    mean_mw   = mean(country_maxwind_mean, na.rm = TRUE),
    min_mw    = min(country_maxwind_mean,  na.rm = TRUE),
    max_mw    = max(country_maxwind_mean,  na.rm = TRUE)
  ) %>%
  print()

maxwind_tercile_colors <- c(
  "Never exposed" = "grey60",
  "Low exposure"  = "steelblue",
  "Mid exposure"  = "goldenrod",
  "High exposure" = "firebrick"
)

irfs_maxwind_tercile <- lp_panel_inter(
  data         = data,
  outcome      = outcome,
  main_var     = "maxwind",
  interact_var = "maxwind_tercile",
  controls     = make_controls("cat1_plus"),
  horizon      = horizon,
  fe           = fe,
  panel_id     = panel_id,
  vcov_formula = vcov_fm
) %>%
  mutate(
    exposure = factor(category,
                      levels = c("Never exposed",
                                 "Low exposure",
                                 "Mid exposure",
                                 "High exposure"))
  )

# Print at h = 5
cat("\n--- Cat1+ × maxwind continuous tercile, h=5 ---\n")
irfs_maxwind_tercile %>%
  filter(horizon == 5) %>%
  select(exposure, irf_mean, se, irf_down, irf_up) %>%
  arrange(exposure) %>%
  as.data.frame() %>%
  print(digits = 4, row.names = FALSE)

fig_maxwind_tercile <- irfs_maxwind_tercile %>%
  ggplot(aes(x = horizon, y = irf_mean,
             color = exposure, fill = exposure)) +
  geom_hline(yintercept = 0, linetype = "dashed",
             linewidth = 0.55, color = "grey40") +
  geom_ribbon(aes(ymin = irf_down, ymax = irf_up),
              alpha = 0.13, color = NA) +
  geom_line(linewidth = 1.05) +
  scale_color_manual(values = maxwind_tercile_colors) +
  scale_fill_manual(values  = maxwind_tercile_colors) +
  scale_x_continuous(breaks = seq(0, horizon, 2)) +
  scale_y_continuous(labels = pct_labels) +
  labs(x     = "Horizon (years)",
       y     = "GDP response",
       color = "Country mean wind speed",
       fill  = "Country mean wind speed") +
  irf_theme +
  theme(legend.position = "top")

fig_maxwind_tercile

ggsave("figures/irf_cat1_maxwind_continuous_tercile.png",
       fig_maxwind_tercile, width = 7, height = 5, dpi = 300)
```

```{r}
# ------------------------------------------------------------------
# IRFs: Cat1+ treatment × country maxwind_95 exposure (binary)
# ------------------------------------------------------------------
data <- data %>%
  group_by(countrycode) %>%
  mutate(country_maxwind95_rate = mean(maxwind_95, na.rm = TRUE)) %>%
  ungroup()

median_rate <- data %>%
  filter(country_maxwind95_rate > 0) %>%
  distinct(countrycode, country_maxwind95_rate) %>%
  pull(country_maxwind95_rate) %>%
  median(na.rm = TRUE)

data <- data %>%
  mutate(
    maxwind95_exposure = case_when(
      country_maxwind95_rate == 0           ~ "Never hit",
      country_maxwind95_rate <= median_rate ~ "Low exposure",
      country_maxwind95_rate >  median_rate ~ "High exposure"
    ),
    maxwind95_exposure = factor(maxwind95_exposure,
                                levels = c("Never hit",
                                           "Low exposure",
                                           "High exposure"))
  )

cat("\n--- Country maxwind_95 binary exposure distribution ---\n")
data %>%
  distinct(countrycode, maxwind95_exposure, country_maxwind95_rate) %>%
  count(maxwind95_exposure) %>%
  print()

data %>%
  distinct(countrycode, maxwind95_exposure, country_maxwind95_rate) %>%
  group_by(maxwind95_exposure) %>%
  summarise(
    n         = n(),
    mean_rate = mean(country_maxwind95_rate, na.rm = TRUE),
    min_rate  = min(country_maxwind95_rate,  na.rm = TRUE),
    max_rate  = max(country_maxwind95_rate,  na.rm = TRUE)
  ) %>%
  print()

exposure_colors <- c(
  "Never hit"     = "#d9d9d9",
  "Low exposure"  = "#9ecae1",
  "High exposure" = "#c0392b"
)

# ------------------------------------------------------------------
# IRF — no legend
# ------------------------------------------------------------------
irfs_maxwind95_exposure <- lp_panel_inter(
  data         = data,
  outcome      = outcome,
  main_var     = "cat1_plus",
  interact_var = "maxwind95_exposure",
  controls     = make_controls("cat1_plus"),
  horizon      = horizon,
  fe           = fe,
  panel_id     = panel_id,
  vcov_formula = vcov_fm
) %>%
  mutate(
    exposure = factor(category,
                      levels = c("Never hit",
                                 "Low exposure",
                                 "High exposure"))
  )

cat("\n--- Cat1+ × maxwind_95 binary exposure, h=5 ---\n")
irfs_maxwind95_exposure %>%
  filter(horizon == 5) %>%
  select(exposure, irf_mean, se, irf_down, irf_up) %>%
  arrange(exposure) %>%
  as.data.frame() %>%
  print(digits = 4, row.names = FALSE)

fig_maxwind95_exposure <- irfs_maxwind95_exposure %>%
  ggplot(aes(x = horizon, y = irf_mean,
             color = exposure, fill = exposure)) +
  geom_hline(yintercept = 0, linetype = "dashed",
             linewidth = 0.55, color = "grey40") +
  geom_ribbon(aes(ymin = irf_down, ymax = irf_up),
              alpha = 0.13, color = NA) +
  geom_line(linewidth = 1.05) +
  scale_color_manual(values = exposure_colors) +
  scale_fill_manual(values  = exposure_colors) +
  scale_x_continuous(breaks = seq(0, horizon, 2)) +
  scale_y_continuous(labels = pct_labels) +
  labs(x     = "Horizon (years)",
       y     = "GDP response",
       color = "Country TC exposure",
       fill  = "Country TC exposure") +
  irf_theme +
  theme(legend.position = "none")   # <-- no legend on IRF

# ------------------------------------------------------------------
# Map — with legend
# ------------------------------------------------------------------
world <- ne_countries(scale = "medium", returnclass = "sf")

map_df_mw95 <- world %>%
  left_join(
    data %>% distinct(countrycode, maxwind95_exposure),
    by = c("adm0_a3" = "countrycode")
  ) %>%
  mutate(
    maxwind95_exposure = as.character(maxwind95_exposure),
    maxwind95_exposure = case_when(
      is.na(maxwind95_exposure) ~ "No data",
      TRUE                      ~ maxwind95_exposure
    ),
    maxwind95_exposure = factor(maxwind95_exposure,
                                levels = c("Never hit",
                                           "Low exposure",
                                           "High exposure",
                                           "No data"))
  ) %>%
  filter(maxwind95_exposure != "No data")

fig_map_mw95 <- ggplot(map_df_mw95) +
  geom_sf(aes(fill = maxwind95_exposure), color = "white", linewidth = 0.15) +
  coord_sf(crs = "+proj=robin") +
  scale_fill_manual(
    values = exposure_colors,
    guide  = guide_legend(
      title          = "Country TC exposure",
      title.position = "top",
      nrow           = 1,
      keywidth       = unit(1.2, "cm"),
      keyheight      = unit(0.4, "cm")
    )
  ) +
  theme_void(base_size = 13) +
  theme(
    legend.position  = "bottom",
    legend.title     = element_text(face = "bold", size = 11),
    legend.text      = element_text(size = 10),
    plot.background  = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA)
  )

# ------------------------------------------------------------------
# Combined: map (left, with legend) + IRF (right, no legend)
# ------------------------------------------------------------------
ggsave("figures/map_maxwind95_binary.png",
       fig_map_mw95, width = 7, height = 5, dpi = 300)

ggsave("figures/irf_cat1_maxwind95_binary.png",
       fig_maxwind95_exposure, width = 7, height = 5, dpi = 300)

fig_combined_mw95 <- fig_map_mw95 + fig_maxwind95_exposure + plot_layout(widths = c(1, 1))

ggsave("figures/combined_maxwind95_binary.png",
       fig_combined_mw95, width = 10, height = 5, dpi = 300)

fig_combined_mw95
```


# State Dependent



```{r}
# ── State-dependent LP: shock × recovery state ───────────────────────────────
# Condition on where the economy sits at time of shock, not on whether
# a prior shock occurred. Follows Ramey & Zubairy (2018).
# Three specs: binary 1yr, binary 2yr, continuous (terciles)

data <- data %>%
  group_by(countrycode) %>%
  mutate(
    # State 1: contracting in t-1
    contracting      = as.integer(lag(gdp_diff, 1, default = 0) < 0),

    # State 2: contracting in both t-1 and t-2 (sustained weakness)
    contracting_2yr  = as.integer(
      lag(gdp_diff, 1, default = 0) < 0 &
      lag(gdp_diff, 2, default = 0) < 0
    ),

    # Continuous: lagged growth standardized
    lagged_growth     = lag(gdp_diff, 1, default = 0),
    lagged_growth_std = lagged_growth / sd(lagged_growth, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(
    growth_tercile = case_when(
      lagged_growth_std <= quantile(lagged_growth_std, 1/3, na.rm = TRUE) ~ "Low growth (bottom tercile)",
      lagged_growth_std <= quantile(lagged_growth_std, 2/3, na.rm = TRUE) ~ "Mid growth",
      TRUE                                                                  ~ "High growth (top tercile)"
    ),
    growth_tercile = factor(growth_tercile,
                            levels = c("Low growth (bottom tercile)",
                                       "Mid growth",
                                       "High growth (top tercile)"))
  )

cat(sprintf(
  "State distribution at shock years:\n  contracting (1yr): %d | expanding: %d\n  contracting (2yr): %d | not: %d\n",
  sum(d_treated$contracting[d_treated$maxwind_95 == 1],      na.rm = TRUE),
  sum(!d_treated$contracting[d_treated$maxwind_95 == 1],     na.rm = TRUE),
  sum(d_treated$contracting_2yr[d_treated$maxwind_95 == 1],  na.rm = TRUE),
  sum(!d_treated$contracting_2yr[d_treated$maxwind_95 == 1], na.rm = TRUE)
))

# ── Shared LP settings ───────────────────────────────────────────────────────
lp_controls <- "l(gdp_diff, 1:2) + l(maxwind_95, 1:2)"
lp_fe       <- "countrycode[year] + countrycode[year2] + region^year"

run_state_irf <- function(interact_var, facet_label) {
  lp_panel_inter(
    data         = data,
    outcome      = outcome,
    main_var     = "maxwind_95",
    interact_var = interact_var,
    controls     = lp_controls,
    horizon      = horizon,
    fe           = lp_fe,
    panel_id     = panel_id,
    vcov_formula = vcov_fm
  ) %>%
    mutate(state = category, facet = facet_label)
}

# ── Spec A: binary — contracting in t-1 ──────────────────────────────────────
irf_state_1yr <- run_state_irf("contracting", "A  Contracting in t\u22121") %>%
  mutate(state = recode(state, "0" = "Expanding", "1" = "Contracting"))

# ── Spec B: binary — contracting in t-1 and t-2 ──────────────────────────────
irf_state_2yr <- run_state_irf("contracting_2yr",
                                "B  Contracting in t\u22121 and t\u22122") %>%
  mutate(state = recode(state, "0" = "Expanding", "1" = "Contracting"))

# ── Spec C: continuous — prior growth tercile ────────────────────────────────
irf_state_cont <- run_state_irf("growth_tercile",
                                 "C  Prior growth tercile (continuous)")

# ── Print summary at h = 5 ───────────────────────────────────────────────────
cat("\n--- State-dependent IRF at h = 5 ---\n")
bind_rows(irf_state_1yr, irf_state_2yr, irf_state_cont) %>%
  filter(horizon == 5) %>%
  select(facet, state, irf_mean, se, irf_down, irf_up) %>%
  arrange(facet, state) %>%
  as.data.frame() %>%
  print(digits = 3, row.names = FALSE)

# ── Combined 3-panel plot ─────────────────────────────────────────────────────
state_colors <- c(
  "Expanding"                   = "steelblue",
  "Contracting"                 = "firebrick",
  "Low growth (bottom tercile)" = "firebrick",
  "Mid growth"                  = "goldenrod3",
  "High growth (top tercile)"   = "steelblue"
)

fig_state <- bind_rows(irf_state_1yr, irf_state_2yr, irf_state_cont) %>%
  mutate(facet = factor(facet, levels = c(
    "A  Contracting in t\u22121",
    "B  Contracting in t\u22121 and t\u22122",
    "C  Prior growth tercile (continuous)"
  ))) %>%
  ggplot(aes(x = horizon, y = irf_mean,
             color = state, fill = state)) +
  geom_ribbon(aes(ymin = irf_down, ymax = irf_up),
              alpha = 0.13, color = NA) +
  geom_line(linewidth = 1.1) +
  geom_hline(yintercept = 0, linetype = "dashed",
             linewidth = 0.55, color = "grey40") +
  facet_wrap(~facet, ncol = 3) +
  scale_color_manual(values = state_colors, name = NULL) +
  scale_fill_manual(values  = state_colors, name = NULL) +
  scale_x_continuous(breaks = seq(0, horizon, 2)) +
  scale_y_continuous(labels = function(x) paste0(round(100 * x, 1), "%")) +
  labs(x = "Horizon (years)", y = "GDP response") +
  irf_theme +
  theme(legend.position = "bottom")

ggsave("figures/irf_state_dependent.png",
       fig_state, width = 13, height = 5, dpi = 300)

fig_state
```

# Compounding

```{r}
# Compound TC effects: how a new severe TC affects recovery depending on prior shock history.

d_treated <- d_treated %>%
  group_by(countrycode) %>%
  mutate(
    # ── Prior shock indicators ───────────────────────────────────────────────
    # Was there a severe TC in the past 1, 2, or 3 years?
    shock_lag1     = lag(maxwind_95, 1, default = 0L),
    shock_lag2     = lag(maxwind_95, 2, default = 0L),
    shock_lag3     = lag(maxwind_95, 3, default = 0L),

    # Binary compound flags
    compound_1yr   = as.integer(maxwind_95 == 1 & shock_lag1 == 1),
    compound_2yr   = as.integer(maxwind_95 == 1 & (shock_lag1 | shock_lag2)),
    compound_3yr   = as.integer(maxwind_95 == 1 & (shock_lag1 | shock_lag2 | shock_lag3)),
    isolated       = as.integer(maxwind_95 == 1 & !shock_lag1 & !shock_lag2 & !shock_lag3),

    # Cumulative exposure: how many severe TCs in the past 3 years?
    prior_shocks_3yr = shock_lag1 + shock_lag2 + shock_lag3,
    prior_shocks_3yr = factor(prior_shocks_3yr,
                              levels = 0:3,
                              labels = c("0 prior", "1 prior", "2 prior", "3 prior")),

    # Time since last shock (years) — NA if never shocked before
    years_since_shock = {
      ys <- rep(NA_real_, n())
      last <- NA_real_
      for (i in seq_along(maxwind_95)) {
        if (!is.na(shock_lag1[i]) && shock_lag1[i] == 1) last <- 0
        else if (!is.na(last)) last <- last + 1
        ys[i] <- last
      }
      ys
    },
    # Bin time since last shock
    recovery_stage = case_when(
      is.na(years_since_shock)    ~ "No prior shock",
      years_since_shock == 0      ~ "Immediate (0y gap)",
      years_since_shock <= 2      ~ "Early recovery (1-2y gap)",
      years_since_shock <= 5      ~ "Mid recovery (3-5y gap)",
      TRUE                        ~ "Full recovery (6y+ gap)"
    ),
    recovery_stage = factor(recovery_stage, levels = c(
      "No prior shock",
      "Immediate (0y gap)",
      "Early recovery (1-2y gap)",
      "Mid recovery (3-5y gap)",
      "Full recovery (6y+ gap)"
    ))
  ) %>%
  ungroup()

cat("── Compound event counts ───────────────────────────────────────────────\n")
cat(sprintf("  Isolated (no shock in t-1 to t-3): %d\n",  sum(d_treated$isolated,     na.rm = TRUE)))
cat(sprintf("  Compound 1yr (shock in t-1):        %d\n",  sum(d_treated$compound_1yr, na.rm = TRUE)))
cat(sprintf("  Compound 2yr (shock in t-1 or t-2): %d\n",  sum(d_treated$compound_2yr, na.rm = TRUE)))
cat(sprintf("  Compound 3yr (shock in t-1 to t-3): %d\n",  sum(d_treated$compound_3yr, na.rm = TRUE)))
cat("────────────────────────────────────────────────────────────────────────\n")

# ── Shared LP settings ───────────────────────────────────────────────────────
lp_controls <- "l(gdp_diff, 1:2) + l(maxwind_95, 1:2)"
lp_fe       <- "countrycode[year] + countrycode[year2] + region^year"

run_inter_irf <- function(interact_var, facet_label) {
  lp_panel_inter(
    data         = d_treated,
    outcome      = outcome,
    main_var     = "maxwind_95",
    interact_var = interact_var,
    controls     = lp_controls,
    horizon      = horizon,
    fe           = lp_fe,
    panel_id     = panel_id,
    vcov_formula = vcov_fm
  ) %>%
    mutate(group = category, facet = facet_label)
}

# ── Spec A: shock × prior shock in t-1 (strict 1yr window) ──────────────────
irf_compound_1yr <- run_inter_irf("shock_lag1", "Prior shock in t\u22121") %>%
  mutate(group = recode(group,
    "0" = "No prior shock",
    "1" = "Prior shock in t\u22121"
  ))

# ── Spec B: shock × cumulative prior shocks (0 / 1 / 2-3) ───────────────────
d_treated <- d_treated %>%
  mutate(
    prior_shocks_bin = case_when(
      prior_shocks_3yr == "0 prior" ~ "0 prior shocks",
      prior_shocks_3yr == "1 prior" ~ "1 prior shock",
      TRUE                          ~ "2-3 prior shocks"
    ),
    prior_shocks_bin = factor(prior_shocks_bin,
                              levels = c("0 prior shocks",
                                         "1 prior shock",
                                         "2-3 prior shocks"))
  )

irf_compound_cum <- run_inter_irf("prior_shocks_bin",
                                   "# prior shocks in t\u22121 to t\u22123")

# ── Spec C: shock × time since last shock (recovery stage) ──────────────────
irf_compound_gap <- run_inter_irf("recovery_stage",
                                   "C  Time since last shock") %>%
  mutate(group = factor(group, levels = c(
    "Immediate (0y gap)",
    "Early recovery (1-2y gap)",
    "Mid recovery (3-5y gap)",
    "Full recovery (6y+ gap)",
        "No prior shock"
  )))

# ── Print summary at h = 5 ──────────────────────────────────────────────────
cat("\n--- Compound TC IRFs at h = 5 ---\n")
bind_rows(irf_compound_1yr, irf_compound_cum, irf_compound_gap) %>%
  filter(horizon == 5) %>%
  select(facet, group, irf_mean, se, irf_down, irf_up) %>%
  arrange(facet, group) %>%
  as.data.frame() %>%
  print(digits = 3, row.names = FALSE)

# ── Plot ─────────────────────────────────────────────────────────────────────
irf_base <- function(df, color_map) {
  ggplot(df, aes(x = horizon, y = irf_mean, color = group, fill = group)) +
    geom_ribbon(aes(ymin = irf_down, ymax = irf_up),
                alpha = 0.13, color = NA) +
    geom_line(linewidth = 1.1) +
    geom_hline(yintercept = 0, linetype = "dashed",
               linewidth = 0.55, color = "grey40") +
    scale_color_manual(values = color_map, name = NULL) +
    scale_fill_manual(values  = color_map, name = NULL) +
    scale_x_continuous(breaks = seq(0, horizon, 2)) +
    scale_y_continuous(labels = function(x) paste0(round(100 * x, 1), "%")) +
    labs(x = "Horizon (years)", y = "GDP response") +
    irf_theme +
    theme(legend.position  = "bottom",
          legend.text      = element_text(size = 10),
          legend.key.width = unit(1.2, "cm")) +
    guides(color = guide_legend(nrow = 2, byrow = TRUE),
           fill  = guide_legend(nrow = 2, byrow = TRUE))
}

# ── Fig A: prior shock in t-1 ────────────────────────────────────────────────
pa <- irf_base(
  irf_compound_1yr,
  c("No prior shock"          = "steelblue",
    "Prior shock in t\u22121" = "firebrick")
) + labs(title = "Prior shock in t\u22121")


# ── Fig B: cumulative prior shocks ───────────────────────────────────────────
pb <- irf_base(
  irf_compound_cum,
  c("0 prior shocks"   = "steelblue",
    "1 prior shock"    = "goldenrod3",
    "2-3 prior shocks" = "firebrick")
) + labs(title = "# prior shocks in t\u22121 to t\u22123")


# ── Fig C: faceted by recovery stage, no legend ──────────────────────────────
colors_gap <- c(
  "Immediate (0y gap)"        = "#8B0000",
  "Early recovery (1-2y gap)" = "firebrick",
  "Mid recovery (3-5y gap)"   = "goldenrod3",
  "Full recovery (6y+ gap)"   = "#2e8b57",
    "No prior shock"            = "steelblue"
)

pc <- irf_base(
  irf_compound_gap %>%
    mutate(group = factor(group, levels = names(colors_gap))),
  colors_gap
) +
  facet_wrap(~group, ncol = 5) +
  labs(title = "IRF by recovery stage at time of new shock") +
  theme(legend.position = "none")

ggsave("figures/irf_compound_stages.png",
       pc, width = 13, height = 5, dpi = 300)

fig_compound <- pa + pb  + plot_layout(ncol = 2)

ggsave("figures/irf_compound_effects.png",
       fig_compound, width = 10, height = 5, dpi = 300)

```



